Practical Machine Learning Course Project
=================================
## Overview
In this project, we will be using the Weight Lifting Exercise Dataset to build a machine learning algorithm to predict activity quality (classe) from activity monitors.

## Load dependencies and set seed
```{r message=FALSE}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
set.seed(1234)
```

## Retrieve and Load datasets
```{r}
# Training dataset	
training_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"

training_file <- "/Users/seanwenweixiong/Documents/Courses/Coursera/Data_Science/8_Practical_Machine_Learning/pml-training.csv"

if(!file.exists(training_file)) {
            download.file(url=training_url, destfile=training_file, method="curl")
            }

trainingset <- read.csv(training_file, na.strings=c("NA", "#DIV/0!", ""))	

# Testing dataset
testing_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

testing_file <-  "/Users/seanwenweixiong/Documents/Courses/Coursera/Data_Science/8_Practical_Machine_Learning/pml-testing.csv"

if(!file.exists(testing_file)) {
            download.file(url=testing_url, destfile=testing_file, method="curl")
            }		

testingset <- read.csv(testing_file, na.strings=c("NA", "#DIV/0!", ""))

# Check dimensions
dim(trainingset); dim(testingset)	
```

## Cleaning datasets
```{r}
# Remove columns with all missing values
trainingset<-trainingset[ ,colSums(is.na(trainingset)) == 0]
testingset <-testingset[ ,colSums(is.na(testingset)) == 0]

# Remove variables irrelevant for building prediction model, specifically the first 7 columns
col_to_rm <- c(1:7)
trainingset   <-trainingset[ ,-col_to_rm]
testingset <-testingset[ ,-col_to_rm]
```

## Cross-validation
- Partition training set into sub-Training and sub-Test to perform cross-validation. 
- This will be performed using random subsampling without replacement.
- The ratio of sub-Training to sub-Test will be 3:1
```{r}
subsamples <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]

# Check dimensions
dim(subTraining); dim(subTesting)
```

## Exploring outcome variable
```{r}
class(subTraining$classe); head(subTraining$classe)

# Frequency for each level
df_count <- data.frame(table(subTraining$classe))
names(df_count) <- c("classe", "count")
df_count

# Proportion of each level
df_prop <- data.frame(table(subTraining$classe)/dim(subTraining)[1]*100)
names(df_prop) <- c("classe", "proportion")
df_prop
```
- The outcome variable (classe) is a factor variable with 5 levels, i.e. A, B, C, D, and E. The proportion of each level is approximately the same.

## Model prediction using decision tree
```{r}
# Fit model
model1 <- rpart(classe ~ ., data=subTraining, method="class")

# Make predictions
prediction1 <- predict(model1, subTesting, type = "class")

# Plot decision tree
	
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)

# Compare predicted values with observed values
confusionMatrix(prediction1, subTesting$classe)

# Accuracy
confusionMatrix(prediction1, subTesting$classe)$overall[1]
```

## Model prediction using random forest
```{r}
# Fit model
model2 <- randomForest(classe ~., data=subTraining, method="class")

# Make predictions
prediction2 <- predict(model2, subTesting, type="class")

# Compare predicted values with observed values
confusionMatrix(prediction2, subTesting$classe)

# Accuracy
confusionMatrix(prediction2, subTesting$classe)$overall[1]
```
- The accuracy of the prediction model using decision tree is 73.9% (95% Cl: 72.7-75.2%). Hence, the expected out-of-sample error here is 26.1%.
- The accuracy of the prediction model using random forest is 99.6% (95% Cl: 99.4-99.8%). Hence, the expected out-of-sample error here is 0.4%.
- Therefore, predictions by random forest is more accurate compared to that of decision tree here.

## Use selected model prediction on testing set
```{r}
prediction3 <- predict(model2, testingset, type="class")
prediction3
```